---
title: "Spatiotemporal DB schema with SSOD data"
author: package author
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spatiotemporal DB schema with SSOD data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 6,
  out.width = "100%",
  fig.align = "center"
)
```
# spatiotemporal
An R package for a spatiotemporal database schema. It has been created to handle weather data from weather stations. An example from GSODR.
See manual for further details.
## Installation from R console
```
## Installiation from Github repo through Gituh token
remotes::install_github("ecor/spatiotemporal")
```
## Usage from R console
```{r, echo=TRUE, results='asis'}
library(spatiotemporal)
help(package="spatiotemporal")
help("get_table")
help("get_ts")
help("get_p")
help("get_variables")
help(db)
```
## Database Schema
Database scheme can be retrieved by this code:
```{r, echo=TRUE, results='asis', fig.show='hold', fig.width=10, fig.height=6, out.width='100%'}
library(spatiotemporal)
library(nomnoml)
library(stringr)
library(data.table)
data(db)
## DATA BASE SCHEMA DESIGN
schema <- lapply(db, FUN=function(t, nfks) {
 #### print(nfks)
  o <- sapply(as.list(t), FUN=function(x) {class(x)})
  o2 <- data.frame(column=names(o), type=as.character(o))
  o2$reference <- as.character(NA)
  iin <- which(o2$column %in% nfks)
  o2$reference[iin] <- o2$column[iin]
  o2$reference[iin] <- str_sub(o2$reference[iin], end=-4)
  return(o2)
}, nfks=sprintf("%s_id", names(db)))
out2 <- list()
for (it in names(schema)) {
  content <- paste(schema[[it]]$column, schema[[it]]$type, sep = ": ", collapse = "||")
  out2[[it]] <- sprintf("[<table>%s| %s]", it, content)
  inn <- which(!is.na(schema[[it]]$reference))
  if (length(inn) > 0) {
    refs <- sprintf("[%s] <- 1..n[%s]", schema[[it]]$reference[inn], it)
    out2[[it]] <- c(out2[[it]], refs)
  }
}
nomnoml(paste(unlist(out2), collapse="\n"), png=TRUE, width=1400, height=700)
```
## Database Population with GSOD data : an example
THis examples shows the direct download data for years from 2015 to 2024 in Morocco. Assiming the same operator, a new project and a new action are created in the respective tables:
```{r, echo=TRUE, results='asis'}
data(db)
##
db0 <- db
##
years <- 2015:2024
country <- "Morocco"
project_name <- country
action_description <- sprintf("%s_%s",country,paste(range(years),collapse="_"))
### new project
project_id <- nrow(db0$project)+1
db$project <- rbind(db$project, data.table::data.table(ID = project_id, name = project_name))
### new action
action_id <- nrow(db0$action)+1
operator_id <- 1
new_action <- data.table::data.table(ID=action_id,description=action_description,timestamptz=Sys.time(),operator_id=operator_id,project_id=project_id)
db$action <- rbind(db$action,new_action)
```
After defing the variables **`years`** and **`country`** , GSOD time series data are reterieved or downloaded as follows (see **`GSODR::get_gsod`**):
```{r, echo=TRUE, results='asis'}
library(GSODR)
weather_ts_data_rds <- sprintf("/home/ecor/local/rpackages/jrc/spatiotemporal/inst/ext_data/%s.rds",action_description)
if (file.exists(weather_ts_data_rds)) {
  weather_ts_data <- readRDS(weather_ts_data_rds)
} else {
  weather_ts_data <- get_GSOD(years = years, country = country)
}
```
New weather stations ara added to **`db$p`** table as follows:
```{r, echo=TRUE, results='asis', fig.show='hold', fig.width=10, fig.height=6, out.width='100%'}
library(terra)
library(sf)
p <- weather_ts_data |>
  dplyr::select(STNID, NAME, CTRY, COUNTRY_NAME, ISO2C, ISO3C, STATE,
                LATITUDE, LONGITUDE, ELEVATION, BEGIN, END) |>
  dplyr::filter(!duplicated(STNID))
prefix <- "GSOD station: data through GSODR R package"
# p$description <- apply(p, 1, function(row) {
#   paste(c(prefix, paste(names(row), row, sep = " : ")), collapse = " ; ")
# })
p$description <- purrr::pmap_chr(p, function(...) {
  row <- list(...)
  paste(c(prefix, paste(names(row), row, sep = " : ")), collapse = " ; ")
})
p <- sf::st_as_sf(
  p,
  coords = c("LONGITUDE", "LATITUDE"),
  crs = st_crs(db$p),
  remove = FALSE
)
####
p$name2 <- sprintf("GSOD_%s (%s - %s)",p$STNID,p$NAME,p$ISO2C)
p$provider_id <- db$provider$ID[db$provider$name=="gsod"]
p$ID <- nrow(db0$p)+1:nrow(p)
db$p <- rbind(db0$p,p[,names(db0$p)])
plet(vect(db$p),cex=20)
```
Variables and attiributes are already defined in db\$variable and b\$attribute tables respectively.
```{r, echo=TRUE, results='asis'}
library(dplyr)
v <- weather_ts_data
v$timestamptz <- as.Date(v$YEARMODA) |> as.POSIXct(tz="GMT")
v$name2 <- sprintf("GSOD_%s (%s - %s)",v$STNID,v$NAME,v$ISO2C)
vp <- db$p |> as.data.table()  |> dplyr::mutate(p_id=ID) |> dplyr::select(p_id,name2)
v <- left_join(v,vp)
ids <- c("timestamptz","p_id")
vv <- v  |> dplyr::select(which(names(v) %in% c(ids,db$variable$name))) |> melt(id.vars=ids)
vv <- db$variable |> mutate(variable_id=ID,variable=name) |> select(variable_id,variable) |> right_join(vv) |> select(-variable)
va <- v  |> dplyr::select(which(names(v) %in% c(ids,db$attribute$name))) |> melt(id.vars=ids) |> mutate(attribute_name=variable,attrvalue=value) |> select(-variable,-value)
##
va <- db$attribute |> mutate(attribute_id=ID,attribute_name=name) |> select(attribute_id,attribute_name,variable_name) |> right_join(va) ##|> select(-variable)
##
va <- db$variable |> mutate(variable_id=ID,variable_name=name) |> select(variable_id,variable_name) |> right_join(va) |> select(-variable_name)
##
###
vv <- full_join(vv,va)
vv$action_id <- action_id
vv$ID <- nrow(db0$ts)+1:nrow(vv)
nts <- names(db$ts)
vv <- vv[,..nts]
###
db$ts <- rbind(db$ts,vv)
tail(db$ts,n=20)
```
## Database Population with SSOD  : an example

Weather station time series can also be retrieved from NOAA's Synoptic Surface Summary of the Day (SSOD, formerly Global Summary of the Day / GSOD), a daily surface-observation dataset maintained by NOAA's National Centers for Environmental Information (NCEI). See the [SSOD dataset page](https://catalog.data.gov/dataset/synoptic-summary-of-the-day-ssod-version-2) for details.
This examples shows the direct download data from GHCN for years from 2025 to 2026 in Morocco. Assiming the same operator, a new project and a new action are created in the respective tables:
```{r, echo=TRUE, results='asis'}
##
db1 <- db
##
country <- "Morocco"
years <- 1900:2026
project_name <- sprintf("ssod_%s",country)
action_description <- sprintf("%s_%s",country,paste(range(years),collapse="_"))
### new project
project_id <- nrow(db1$project)+1
db$project <- rbind(db1$project, data.table::data.table(ID = project_id, name = project_name))
### new action
action_id <- nrow(db1$action)+1
operator_id <- 1
new_action <- data.table::data.table(ID=action_id,description=action_description,timestamptz=Sys.time(),operator_id=operator_id,project_id=project_id)
db$action <- rbind(db$action,new_action)
```
Direct download of the data throwgh the following script:
```{r,echo=TRUE,results='asis'}
## ---------------------------------------------------------------------------
## R script: download SSOD v2 data for ALL available Morocco weather
## stations, years 2000-2026, using link #1 ("by-year" access point)
##
##   https://www.ncei.noaa.gov/oa/synoptic-summary-of-the-day/v2/access/by-year/
##
## Morocco's FIPS country code in this dataset is "MO" (confirmed against
## Moroccan airport ICAO codes present in the bucket, e.g. "GMMN" =
## Mohammed V Intl, Casablanca).
##
## The list of Moroccan stations is NOT hardcoded: for every year, the
## script queries the underlying S3-compatible ListObjectsV2 endpoint
## with prefix "v2/access/by-year/<year>/csv/SSOD_MO" and downloads
## whatever station files come back. This automatically adapts to
## stations being added or dropped over time (verified: 25 Moroccan
## stations in 2000 vs. 31 in 2020 - the station list does change).
##
## NOTE: this script makes one listing request + several dozen file
## downloads PER YEAR, so downloading the full 2000-2026 range can take
## a while and will create several hundred small CSV files locally.
## ---------------------------------------------------------------------------
pkgs <- c("httr", "xml2")
to_install <- pkgs[!sapply(pkgs, requireNamespace, quietly = TRUE)]
if (length(to_install) > 0) install.packages(to_install)
library(httr)
library(xml2)
library(countrycode)
# --- parameters: change these to target a different country/year range ----
country_prefix <- countrycode::countrycode(country, "country.name", "fips")       # Morocco FIPS country code

pause_seconds  <- 0.1         # small delay between downloads, be nice to the server
base_url <- "https://www.ncei.noaa.gov/oa/synoptic-summary-of-the-day/"
###dir.create("ssod_data", showWarnings = FALSE)
#' List every SSOD "by-year" file key for a given year and country prefix
#' (handles pagination, in case a country ever has more than 1000 stations)
list_country_files_for_year <- function(year, country_prefix) {
  prefix <- sprintf("v2/access/by-year/%d/csv/SSOD_%s", year, country_prefix)
  keys <- character()
  token <- NULL
  repeat {
    query <- list(`list-type` = "2", prefix = prefix)
    if (!is.null(token)) query$`continuation-token` <- token
    resp <- GET(base_url, query = query)
    stop_for_status(resp)
    doc <- read_xml(content(resp, as = "text", encoding = "UTF-8"))
    xml_ns_strip(doc)   # strip the S3 namespace to simplify XPath queries
    found <- xml_text(xml_find_all(doc, ".//Contents/Key"))
    keys <- c(keys, found)
    is_truncated <- xml_text(xml_find_first(doc, ".//IsTruncated"))
    if (identical(is_truncated, "true")) {
      token <- xml_text(xml_find_first(doc, ".//NextContinuationToken"))
    } else {
      break
    }
  }
  keys
}
#' Download one file (S3 key) into ssod_data/<year>/<filename>.csv and
#' read it into a data.frame; returns NULL (with a warning) on failure
download_one <- function(key, year,wpath,overwrite=FALSE) {
  file_url <- paste0(base_url, key)
  dest_dir <- file.path(wpath,"/ssod_data", year)
  dir.create(dest_dir, showWarnings = FALSE, recursive = TRUE)
  dest_file <- file.path(dest_dir, basename(key))
  if ((!file.exists(dest_file)) | overwrite) {
  resp <- tryCatch(GET(file_url), error = function(e) NULL)
  if (is.null(resp) || http_error(resp)) {
    warning(sprintf("Failed to download %s - skipping", key))
    return(NULL)
  }
  writeBin(content(resp, as = "raw"), dest_file)
  }
  tryCatch(read.csv(dest_file, stringsAsFactors = FALSE),
           error = function(e) {
             warning(sprintf("Failed to parse %s - skipping", key))
             NULL
           })
}
#' rbind a list of data.frames whose columns may not perfectly match
#' across years (missing columns are filled with NA)
rbind_fill <- function(df_list) {
  df_list <- Filter(Negate(is.null), df_list)
  if (length(df_list) == 0) return(data.frame())
  all_cols <- unique(unlist(lapply(df_list, names)))
  df_list <- lapply(df_list, function(df) {
    missing_cols <- setdiff(all_cols, names(df))
    for (col in missing_cols) df[[col]] <- NA
    df[all_cols]
  })
  do.call(rbind, df_list)
}
# --- main loop: discover + download every Morocco station, every year -----
all_data        <- list()
station_summary <- data.frame(year = integer(), n_stations = integer())
wpath <- "/home/ecor/local/data/climate/ssod"
for (year in years) {
  cat("\n==== Year", year, "====\n")
  keys <- list_country_files_for_year(year, country_prefix)
  cat("Found", length(keys), "country station file(s)\n")
  station_summary <- rbind(station_summary,
                            data.frame(year = year, n_stations = length(keys)))
  for (key in keys) {
    cat("  Downloading:", basename(key), "\n")
    df <- download_one(key, year,wpath)
    if (!is.null(df)) {
      all_data[[length(all_data) + 1]] <- df
    }
    if (pause_seconds > 0) Sys.sleep(pause_seconds)
  }
}
# --- combine everything into one data.frame ---------------------------------
combined <- rbind_fill(all_data)
combined[combined==-9999.9] <- NA
cat("\n---------------------------------------------\n")
cat("Years processed:", paste(range(years), collapse = "-"), "\n")
cat("Total rows downloaded across all stations/years:", nrow(combined), "\n")
cat("\nNumber of Morocco stations found per year:\n")
print(station_summary)
```
SSOD column name correspond to the respective GSOD names.
```{r,echo=TRUE,results='asis'}
# Lookup table mapping the `combined` data frame's column names
# to the standard NOAA GSOD/SSOD field names and their meaning
gsod_lookup <- data.frame(
  combined_column = c(
    "STATION",
    "Station_name",
    "DATE",
    "Year",
    "Month",
    "Day",
    "LATITUDE",
    "LONGITUDE",
    "ELEVATION",
    "mean_temperature",
    "mean_temperature_Measurement_Code",
    "mean_dew_point_temperature",
    "mean_dew_point_temperature_Measurement_Code",
    "mean_sea_level_pressure",
    "mean_sea_level_pressure_Measurement_Code",
    "mean_station_level_pressure",
    "mean_station_level_pressure_Measurement_Code",
    "mean_visibility",
    "mean_visibility_Measurement_Code",
    "mean_wind_speed",
    "mean_wind_speed_Measurement_Code",
    "max_wind_speed",
    "max_wind_gust",
    "max_temperature",
    "max_temperature_Measurement_Code",
    "min_temperature",
    "min_temperature_Measurement_Code",
    "total_precipitation",
    "total_precipitation_Measurement_Code",
    "snow_depth",
    "observed_pres_wx"
  ),
  gsod_field = c(
    "STATION",
    "NAME",
    "DATE",
    NA,                    # derived from DATE, not a raw GSOD field
    NA,                    # derived from DATE, not a raw GSOD field
    NA,                    # derived from DATE, not a raw GSOD field
    "LATITUDE",
    "LONGITUDE",
    "ELEVATION",
    "TEMP",
    "TEMP_ATTRIBUTES",
    "DEWP",
    "DEWP_ATTRIBUTES",
    "SLP",
    "SLP_ATTRIBUTES",
    "STP",
    "STP_ATTRIBUTES",
    "VISIB",
    "VISIB_ATTRIBUTES",
    "WDSP",
    "WDSP_ATTRIBUTES",
    "MXSPD",
    "GUST",
    "MAX",
    "MAX_ATTRIBUTES",
    "MIN",
    "MIN_ATTRIBUTES",
    "PRCP",
    "PRCP_ATTRIBUTES",
    "SNDP",
    "FRSHTT"
  ),
  description = c(
    "Station identifier (USAF+WBAN)",
    "Station name",
    "Observation date",
    "Year (derived from DATE)",
    "Month (derived from DATE)",
    "Day (derived from DATE)",
    "Station latitude, decimal degrees",
    "Station longitude, decimal degrees",
    "Station elevation, meters",
    "Daily mean temperature",
    "Number of hourly obs (0-24) used for the mean temperature",
    "Daily mean dew point temperature",
    "Number of hourly obs (0-24) used for the mean dew point",
    "Daily mean sea-level pressure",
    "Number of hourly obs (0-24) used for the mean SLP",
    "Daily mean station-level pressure",
    "Number of hourly obs (0-24) used for the mean STP",
    "Daily mean visibility",
    "Number of hourly obs (0-24) used for the mean visibility",
    "Daily mean wind speed",
    "Number of hourly obs (0-24) used for the mean wind speed",
    "Maximum sustained wind speed (no attribute field)",
    "Maximum wind gust (no attribute field)",
    "Daily maximum temperature",
    "Flag: blank = reported synoptic max, '*' = derived from hourly TEMP",
    "Daily minimum temperature",
    "Flag: blank = reported synoptic min, '*' = derived from hourly TEMP",
    "Daily total precipitation",
    "Flag A-I: accumulation basis (6h/12h/24h reports, or 'I' = no data)",
    "Snow depth (no attribute field)",
    "6-digit FRSHTT occurrence flag: Fog/Rain-Drizzle/Snow-Ice/Hail/Thunder/Tornado"
  ),
  stringsAsFactors = FALSE
)
inn <- which(!is.na(gsod_lookup$gsod_field))
combined2 <- combined[,inn]
weather_ts_data <- combined2
names(weather_ts_data) <- gsod_lookup$gsod_field[inn]
```
New weather stations from SSOD are added to **`db$p`** table as follows. Unlike GSODR, the SSOD download has no separate station-history file, so `CTRY`/`COUNTRY_NAME`/`ISO2C`/`ISO3C` are filled from the `country`/`country_prefix` values fixed above (this download targets a single country), `STATE` is left `NA`, and `BEGIN`/`END` are taken as the observed date range per station in this download rather than an official period-of-record:
```{r, echo=TRUE, results='asis', fig.show='hold', fig.width=10, fig.height=6, out.width='100%'}
library(terra)
library(sf)
library(countrycode)
weather_ts_data$DATE <- as.Date(as.character(weather_ts_data$DATE))
## one row per station: coordinates/metadata (assumed constant per station)
## plus the observed date range for this station in this download
p <- weather_ts_data |>
  dplyr::group_by(STATION) |>
  dplyr::summarise(
    NAME      = dplyr::first(NAME),
    LATITUDE  = dplyr::first(LATITUDE),
    LONGITUDE = dplyr::first(LONGITUDE),
    ELEVATION = dplyr::first(ELEVATION),
    BEGIN     = min(DATE, na.rm = TRUE),
    END       = max(DATE, na.rm = TRUE),
    .groups   = "drop"
  ) |>
  dplyr::rename(STNID = STATION) |>
  dplyr::filter(!duplicated(STNID)) |>
  as.data.frame()
## SSOD rows don't carry per-station CTRY/COUNTRY_NAME/ISO2C/ISO3C the way
## GSODR does; since this download targeted a single country, fill them in
p$CTRY         <- country_prefix
p$COUNTRY_NAME <- country
p$ISO2C        <- countrycode::countrycode(country, "country.name", "iso2c")
p$ISO3C        <- countrycode::countrycode(country, "country.name", "iso3c")
p$STATE        <- as.character(NA)
prefix <- "SSOD station: data through NCEI Synoptic Summary of the Day v2"
p$description <- purrr::pmap_chr(
  p[, c("STNID", "NAME", "CTRY", "COUNTRY_NAME", "ISO2C", "ISO3C", "STATE",
        "LATITUDE", "LONGITUDE", "ELEVATION", "BEGIN", "END")],
  function(...) {
    row <- list(...)
    paste(c(prefix, paste(names(row), row, sep = " : ")), collapse = " ; ")
  }
)
p <- sf::st_as_sf(
  p,
  coords = c("LONGITUDE", "LATITUDE"),
  crs = st_crs(db$p),
  remove = FALSE
)
####
p$name2 <- sprintf("SSOD_%s (%s - %s)", p$STNID, p$NAME, p$ISO2C)
## assumes a "ssod" row already exists in db$provider (as "gsod" does);
## created here if missing so the lookup below never returns NA
if (!("ssod" %in% db$provider$name)) {
  db$provider <- rbind(
    db$provider,
    data.table::data.table(ID = nrow(db$provider) + 1, name = "ssod")
  )
}
p$provider_id <- db$provider$ID[db$provider$name == "ssod"]
p$ID <- nrow(db1$p) + 1:nrow(p)
db$p <- rbind(db1$p, p[, names(db1$p)])
plet(vect(db$p), cex = 20)
```
SSOD also reports two fields that share their names with the standard GSOD schema: snow depth (`SNDP`) and a weather-occurrence flag (`FRSHTT`) — both already documented in `gsod_lookup` above. Neither has an entry yet in **`db$variable`**, so before `db$ts` can be populated for SSOD they are added here, using `gsod_lookup$description` for their `description` and associating them with the `gsod` provider (they are GSOD-standard field names, not an SSOD-specific concept), the same way `MXSPD`/`GUST` already exist with no companion `db$attribute` row (they have no `_ATTRIBUTES`/measurement-code column, so they need no attribute entry):
```{r, echo=TRUE, results='asis'}
new_variable_names <- c("SNDP", "FRSHTT")
new_variable_names <- setdiff(new_variable_names, db$variable$name)
if (length(new_variable_names) > 0) {
  new_variables <- data.table::data.table(
    ID          = nrow(db$variable) + seq_along(new_variable_names),
    name        = new_variable_names,
    description = gsod_lookup$description[match(new_variable_names, gsod_lookup$gsod_field)],
    provider_id = db$provider$ID[db$provider$name == "gsod"]
  )
  ## fill = TRUE: safely adds the rows even if db$variable carries extra
  ## columns beyond ID/name/description/provider_id, leaving those NA here
  ## rather than guessing at a schema that isn't shown in this vignette
  db$variable <- rbind(db$variable, new_variables, fill = TRUE)
}
tail(db$variable)
```
Finally, the retrieved SSOD time series data are inserted into the **`db$ts`** table, following the same melt/join procedure used for the GSOD data above. `weather_ts_data` now carries GSOD-standard field names (`STATION`, `NAME`, `DATE`, `TEMP`, ..., `SNDP`, `FRSHTT`) but, unlike the raw GSODR feed, no `ISO2C` column, so rows are matched to their station via `STNID`/`STATION` against the `ssod` rows of **`db$p`** rather than via the `name2` label used for GSOD; everything else — melting the variable columns against `db$variable`, melting the `_ATTRIBUTES` columns against `db$attribute`, and tagging the result with the current SSOD `action_id` — mirrors the GSOD chunk exactly:
```{r, echo=TRUE, results='asis'}
library(dplyr)
v <- weather_ts_data
v$timestamptz <- as.POSIXct(v$DATE, tz = "GMT")
## match rows to their station by STNID/STATION, restricted to the ssod
## provider (weather_ts_data has no ISO2C/name2 to join on, unlike GSOD)
vp <- db$p |> as.data.table() |>
  dplyr::filter(provider_id == db$provider$ID[db$provider$name == "ssod"]) |>
  dplyr::mutate(p_id = ID) 
prefix <- "SSOD_"
lpp <- str_length(prefix)
vp$code <- str_split(vp$name2,"[ (]") |> sapply(FUN=function(x){x[[1]]}) 
v$code <- paste0(prefix,v$STATION)


#|>
#  dplyr::select(p_id, STNID)
v <- dplyr::left_join(v, vp) ###, by = c("STATION" = "STNID"))
ids <- c("timestamptz","p_id")
vv <- v  |> dplyr::select(which(names(v) %in% c(ids,db$variable$name))) |> as.data.table() |>  melt(id.vars=ids)
vv <- db$variable |> mutate(variable_id=ID,variable=name) |> select(variable_id,variable) |> right_join(vv) |> select(-variable)
va <- v  |> dplyr::select(which(names(v) %in% c(ids,db$attribute$name))) |> as.data.table() |> melt(id.vars=ids) |> mutate(attribute_name=variable,attrvalue=value) |> select(-variable,-value)
##
va <- db$attribute |> mutate(attribute_id=ID,attribute_name=name) |> select(attribute_id,attribute_name,variable_name) |> right_join(va) ##|> select(-variable)
##
va <- db$variable |> mutate(variable_id=ID,variable_name=name) |> select(variable_id,variable_name) |> right_join(va) |> select(-variable_name)
##
###
vv <- full_join(vv,va)
vv$action_id <- action_id
vv$ID <- nrow(db1$ts)+1:nrow(vv)
nts <- names(db$ts)
vv <- vv[,..nts]
###
db$ts <- rbind(db$ts,vv)
tail(db$ts,n=20)
```




## Interactive PRCP explorer: click a station to see its time series

The chunk below needs no Shiny server: `crosstalk` links the `leaflet`
map of stations to a `plotly` line chart of daily precipitation
(**PRCP**), purely client-side, so it works in a static
`rmarkdown::html_vignette`. Each station keeps the same color on the
map and in the chart (e.g. Casablanca's marker and Casablanca's PRCP
line share one color) but there is no name legend; hover a marker for
its name, or click a marker (or a line) to highlight just that
station's PRCP series — double-click the chart to reset.

```{r, echo=TRUE, message=FALSE, warning=FALSE,fig.width=100}
library(leaflet)
library(plotly)
library(crosstalk)
library(dplyr)

max_stations <- 1 #

## PRCP time series for every station just added from GSOD, so we can
## pick the `max_stations` that actually HAVE precipitation data
## (some GSOD stations report no PRCP at all, which is why a naive
## "first 15 by ID" pick can end up with empty/invisible lines)
prcp_all <- get_ts(
  x = db,
  p_id = unique(p$ID),
  get_variable_args = list(name = "PRCP")
) |>
  dplyr::filter(!is.na(value))

demo_ids <- prcp_all |>
  dplyr::count(p_id, name = "n_obs") |>
  dplyr::arrange(dplyr::desc(n_obs)) |>
  dplyr::slice_head(n = max_stations) |>
  dplyr::pull(p_id)

prcp <- prcp_all |> dplyr::filter(p_id %in% demo_ids)

## Attach a human-readable station label (character keys, to avoid a
## silent numeric/character join mismatch dropping the match)
station_lookup <- sf::st_drop_geometry(db$p)[, c("ID", "name2")]
station_lookup$ID <- as.character(station_lookup$ID)
prcp$p_id <- as.character(prcp$p_id)

prcp <- prcp |> dplyr::left_join(station_lookup, by = c("p_id" = "ID"))

stations_sf <- db$p[as.character(db$p$ID) %in% as.character(demo_ids), ]
stations_sf$id_chr <- as.character(stations_sf$ID)

## One color per station (by name), reused for both the map marker
## and its matching PRCP line
station_colors <- grDevices::hcl.colors(length(unique(stations_sf$name2)), palette = "Dark 3")
names(station_colors) <- unique(stations_sf$name2)
stations_sf$color <- station_colors[stations_sf$name2]

## A shared crosstalk key (station ID, as character on both sides) links
## the map and the chart
sd_stations <- crosstalk::SharedData$new(stations_sf, key = ~id_chr, group = "prcp_explorer")
sd_prcp     <- crosstalk::SharedData$new(prcp,        key = ~p_id,   group = "prcp_explorer")

station_map <- leaflet::leaflet(sd_stations, height = 450) |>
  leaflet::addProviderTiles(leaflet::providers$OpenTopoMap) |>
  leaflet::addCircleMarkers(
    radius = 7, stroke = TRUE, color = "#333333", weight = 1,
    fillColor = ~color, fillOpacity = 0.9,
    label = ~name2
  )

ts_plot <- plotly::plot_ly(
  sd_prcp,
  x = ~timestamptz, y = ~value,
  color = ~name2, colors = station_colors,
  height = 450,
  hoverinfo = "text",
  text = ~sprintf("%s<br>%s: %.1f mm", name2, format(timestamptz, "%Y-%m-%d"), value)
) |>
  plotly::add_lines(line = list(width = 1)) |>
  plotly::layout(
    showlegend = FALSE,
    xaxis = list(title = "", rangeslider = list(visible = TRUE)),
    yaxis = list(title = "PRCP (mm)")
  ) |>
  plotly::highlight(
    on = "plotly_click", off = "plotly_doubleclick",
    opacityDim = 0.05, selectize = FALSE, persistent = FALSE
  )

crosstalk::bscols(widths = c(5, 7), station_map, ts_plot)
```
